---
title: "Collection (La Junta)"
author: "Connor Krenzer"
date: "3/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup-packages, include = FALSE, eval = TRUE}

knitr::opts_chunk$set(echo = TRUE,
                      include = TRUE,
                      results = "hide",
                      error = FALSE,
                      warning = FALSE,
                      message = FALSE)

# Packages
if(!require(pacman)) install.packages("pacman")
pacman::p_load(rvest, stringr, tidyr, readr, dplyr, lubridate, clock, ggplot2, htmltools, htmlwidgets)

# The collection() function
source("Collection/functions/collection().R")

```


# Implementation

I have been changing the URLs around to find different data, and the market reports seem to start around ID == 6700. The max URL at the time of writing is around 14000. The data begins in 2016.

```{r urls-execute, eval = TRUE}

# This parses together different market report IDs and stores them in a vector
# (useful for when you do not know the URLs beforehand)
#urls <- paste0("http://www.winterlivestock.com/lajunta.php?reportID=", 6680:14000, "#marketreport")


#the most recent market report, without the permalink
#NOTE: the URL column will not contain the market report number!!!
#urls <- "http://www.winterlivestock.com/lajunta.php"


# This contains all the La Junta market report URLs for easy access
# (recommended if you plan to re-collect the data from scratch--just
# do not forget the possibility that some market reports could have
# been missed, if you really think it necessary to recollect this data)
urls <- read_lines("Collection/La Junta URLs.txt")


# the last-used market report URL
urls <- "http://www.winterlivestock.com/lajunta.php?reportID=13741#marketreport"


# This line is used to automate the update process through RPAs
urls <- read_lines("Collection/current_url_Lajunta.txt")

# No need to save the output--this function writes to a csv file...
collection(urls = urls)

```






# Remove File

If you would like to remove the csv file on your computer when updating data (meaning you would like to start from scratch), you can run the below chunk to remove the file.

```{r remove csv, eval = FALSE}

if(file.exists("La Junta Market Reports.csv")){
  file.remove("La Junta Market Reports.csv")
}

```



# Cleaning

The final step before we are ready for analysis is data cleaning. There are two primary goals in the cleaning script. First, we want to add column names to the csv. The second goal is to organize cattle into a smaller number of distinct groups.

Goal 1 is a piece of cake, but goal 2 is easier said than done. You can analyze some of my judgment calls and decide for yourself what needs to be done to group the different types of livestock.

Do you think that prices should be adjusted for inflation at some point in the future?

If you have trouble reading this csv file, click on Tools \> Global Options \> RMarkdown, then set 'Evaluate chunks in directory:' to Project.

A copy of the csv file will be created so we don't have to rerun the collection() function should the cleaning chunk need updating.

```{r cleaning, eval = TRUE}

# Step 0: save a copy of the uncleaned file if we don't already have one!
if(!file.exists("La Junta Market Reports (before cleaning).csv")){
  
  # "La Junta Market Reports.csv" is unclean if this line executes
  file.copy(from = "La Junta Market Reports.csv", to = "La Junta Market Reports (before cleaning).csv", overwrite = TRUE)
  
}

# Reads in the uncleaned version of the csv, which should be a copy
# of "La Junta Market Reports.csv" if you are 'starting fresh'
lajunta <- readr::read_csv("La Junta Market Reports (before cleaning).csv", 
                           col_names = FALSE)




# Step 1: adding a header to the csv --------------------------
# Adding in the column names
lajunta <- lajunta %>% 
  rename("Date" = X1,
         "Buyer" = X2,
         "Quantity" = X3,
         "Type" = X4,
         "Weight" = X5,
         "Price" = X6,
         "URL" = X7)

# Putting the entries in chronological order,
# first changing the data type to a Date:
lajunta$Date <- lubridate::mdy(lajunta$Date)





# Step 2: reproductive types ---------------------------------
# Making a column assigning cattle reproductive status
# (removing values with 's' as the last letter before this step)
lajunta <- lajunta %>% 
  mutate(Type = str_remove_all(Type, "s$")) %>% 
  mutate(Reprod = str_extract(Type, "hfr$|str$|bull$|cow$|heifer$|steer$|pair$|bow$|hrf$|hr$"),
         .before = 7)

lajunta <- lajunta %>% 
  mutate(Reprod = Reprod %>% 
           str_replace_all("pair", "cow") %>% #replacing 'pair' with cow. It seemed reasonable...
           str_replace_all("bow", "cow") %>%  #replacing a typo with the intended value
           str_replace_all("hrf|hr", "hfr") %>% #replacing typos with the intended values
           str_replace_all("heifer", "hfr") %>% #shortening name
           str_replace_all("steer", "str")) #shortening name















# Step 3: categorizing cattle ---------------------------------
# Removing the plural of the type (Ex. "black cows" becomes "black cow")
lajunta$Type <- str_remove(lajunta$Type, "s$")












# The bulk of the edits
lajunta$Type <- lajunta %>% 
  dplyr::select(Type) %>% 
  unlist() %>% 
  str_remove_all("\\.|,") %>%                     # removing punctuation
  str_remove_all("\\sx.*$|\\s[^\\s]*$") %>%  # removing " x hfr", " hfr", " x str", etc.
  str_replace_all("angus", "ang") %>% 
  str_remove_all("sim-") %>%                      # removing the sim- from sim-angus
  str_replace_all("sim\\s+ang", "ang") %>%        # removing "sim " from "sim angus"
  str_replace_all("\\s*sim", " ang") %>%          # replacing " sim" with " ang"
  str_replace_all("beefmaster|beefmstr", "bfmstr") %>%     # replacing "beefmaster" with "bfmstr"
  str_replace_all("shorthorn|shthrn", "sthrn") %>%# replacing "shorthorn" with "sthrn"
  str_replace_all("clr longhorn", "lnhrn") %>%    # replacing "longhorn" with "lnhrn" 
  str_replace_all("limousin", "lim") %>% 
  str_replace_all("limo", "lim") %>%              # replacing "limo" with "lim"
  str_replace_all(".*\\s+lim|lim flex", "lim") %>%# combining all different types of limousins 
  str_replace_all("gelbvieh|gelb|gelvieh", "gel") %>% 
  str_replace_all("hereford", "here") %>% 
  str_replace_all("charolais", "char") %>% 
  str_replace_all("brahman", "brah") %>%
  str_remove_all("wf-|bwf-|rwf-") %>%             # removing face types from the data when there is other info
  str_remove_all("\\s*&") %>%                     # removing ampersands
  str_replace_all("\\s?blk\\s?", "black") %>%     # replacing abbreviation "blk" with "black"
  str_replace_all("red black", "black red") %>%   # reordering "red and black" to "black and red"
  str_replace_all(".*-.*|mixed", "mix") %>%       # replaces any named cross-breed with "mix"
  str_replace_all("ang char|char ang", "mix") %>% # replaces "char ang" with "mix" explicitly 
  str_replace_all("balancer|bal|bclr", "mix") %>% # replaces "balancer" with "mix", since it's a hybrid
  str_replace_all("santa gert", "mix") %>%        # replaces "santa gert" with "mix", since it's a hybrid
  str_replace_all("stabilizer|stab", "mix") %>%   # replaces "stabilizer" with "mix", since it's a hybrid 
  str_replace_all("barzona", "mix") %>%           # replaces "barzona" with "mix", since it's a hybrid
  str_remove_all("\\s+wf|\\s+bwf|\\s+rwf") %>%    # removing face types at after the name
  str_replace_all("^wf$", "bwf") %>%              # recoding "wf" as "bwf" because it performs similarly and has few observations
  str_remove_all("wf\\s+|bwf\\s+|rwf\\s+") %>%    # removing face types before the name
  #str_replace_all("here|brah", "red") %>%         # replacing "here" and "brah" with "red" due to few observations and similar weight (Since the sample size increased, these cattle can now be kept)
  str_replace_all("black\\s+(ang)", "\\1") %>%    # removing color starts on this line
  str_replace_all("black\\s+(lim)", "\\1") %>% 
  str_replace_all("black\\s+(gel)", "\\1") %>% 
  str_replace_all("black\\s+(here)", "\\1") %>% 
  str_replace_all("black\\s+(char)", "\\1") %>% 
  str_replace_all("red\\s+(ang)", "\\1") %>% 
  str_replace_all("red\\s+(lim)", "\\1") %>% 
  str_replace_all("red\\s+(gel)", "\\1") %>% 
  str_replace_all("red\\s+(here)", "\\1") %>% 
  str_replace_all("red\\s+(char)", "\\1") %>%     # removing color ends on this line
  str_squish() %>% 
  str_trim()





#combining all colors into a new category--"clr" (optional)
lajunta$Type <- lajunta %>% 
  dplyr::select(Type) %>% 
  unlist() %>% 
  str_replace_all("black red|black|balck|red|gray|grey|roan|brown|brwn|bwn", "clr") %>%   # "black red," "black," and "red" are all colors
  str_replace_all("bwf|rwf|wf|spot", "face") %>%
  str_replace_all("face", "clr")                  # remove this line if you want to stop pooling the face group with the color group








# Miscellaneous Corrections -----------------------------------
# This correction of the Type column is done after finding errors explicitly. I am 'hard-coding' in the correct values. I checked the distinct Type values after running the above lines in this section, then fixed the remaining values to fit into a few categories. An admittedly lazy way of doing things, but the file is small--and I'm not a software engineer!
#Note: many of these operations assume that cross-breeds have already been assigned "mix"
lajunta$Type <- lajunta %>%
  dplyr::select(Type) %>%
  unlist() %>%
  str_remove_all("x ") %>%                  # removing stray 'x' characters
  str_replace_all("weaned", "clr") %>%      # Weaned cattle isn't a type, so it gets pooled in with "clr"
  str_replace_all("sal", "clr") %>%         # I don't know what type of cattle "sal" is, so it gets pooled in with "clr"
  str_replace_all("lh", "clr") %>%          # I don't know what type of cattle "lh" is, so it gets pooled in with "clr"
  str_replace_all("brnd", "clr") %>%        # I don't know what type of cattle "brnd" is, so it gets pooled in with "clr"
  str_replace_all("clrrod", "clr") %>%      # I don't know what type of cattle "clrrod" is (before "clr" was appended to the name in an earlier operation), so it gets pooled in with "clr"
  str_replace_all("mis", "clr") %>%         # I don't know what type of cattle "mis" is, so it gets pooled in with "clr"
  str_replace_all(".*char$", "char") %>%    # removing all characters preceding "char"
  str_replace_all("chr|char.*", "char") %>% # replacing "chr" with "char"
  str_replace_all(".*ang$", "ang") %>%      # removing all characters preceding "ang"
  str_replace_all("angmental", "ang") %>%   # replacing "angmental" with "ang"
  str_replace_all(".*maine$", "maine") %>%  # removing all characters preceding "maine"
  str_replace_all("maine", "me") %>%        # replacing "maine" with "me"
  str_remove_all("irish") %>%               # removing the word "irish" anywhere it appears (usually accompanying 'red')
  str_replace_all(".*clr$", "clr") %>%      # removing all characters preceding "clr" (this can only be done after the cases for more specific types [like char] have been addressed)
  str_replace_all("hererord", "here") %>%  # Fixing Errors
  str_replace_all("ang strs", "ang") %>% 
  str_replace_all("mixck", "mix")          # End of Fixing Errors


# After doing more testing it would appear that the remaining breeds do not have a large enough sample to be worth using. All cattle types not in the below list will be assigned to "clr":
types <- c("clr", "ang", "mix", "char", "lim", "gel", "here", "mot")

lajunta$Type[which(!lajunta$Type %in% types)] <- "clr"



# We now have eight categories in the "Type" variable














# Step 4: Writing to file -------------------------------------
write_csv(x = lajunta,
          file = "La Junta Market Reports.csv",
          append = F,
          col_names = T)


# Writing the URLs to a file
write_lines(x = unique(lajunta$URL),
            file = "Collection/La Junta URLs.txt",
            append = F)

```












# Testing


Of course, we want to make sure everything looks good, no? This section identifies problem areas in the file.

```{r testing, eval = FALSE}
#lajunta <- read_csv("https://raw.githubusercontent.com/Ckrenzer/Winter-Livestock-Data/main/La%20Junta%20Market%20Reports.csv")

lajunta <- read_csv("La Junta Market Reports.csv")



# finding market reports using the most recent market report instead of the permalink.
lajunta %>% 
  filter(URL == "http://www.winterlivestock.com/lajunta.php") %>% 
  as.data.frame()




# finds counts of all types--good for finding how many types are in the data
count(lajunta, Type) %>% 
  arrange(desc(n)) %>% 
  as.data.frame()

# finds counts of all reproductive statuses--there should be five: bull, cow, steer, heifer, and NA (if there are missing values).
count(lajunta, Reprod) %>% 
  arrange(desc(n)) %>% 
  as.data.frame()

# How many sales were there on a given day?
lajunta %>% 
  count(Date)

# Taking a look at the newest data
tail(lajunta)


# MISSING VALUES _-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-_-

# relevant info when imputing values:
lajunta %>% 
  group_by(Reprod) %>% 
  dplyr::summarize("Median price" = median(Price, na.rm = T),
                   "Median weight" = median(Weight, na.rm = T),
                   "Median quantity" = median(Quantity, na.rm = T),
                   "Mean Price" = mean(Price, na.rm = T),
                   "Mean weight" = mean(Weight, na.rm = T),
                   "Mean quantity" = mean(Quantity))
# KNN is probably the best imputation method for this data



# DATE -__-__-__-__-__-__-__-__-__-__-
# how many missing values are in the Date column?
lajunta %>% 
  filter(is.na(Date)) %>% 
  slice_sample(n = 20)


# BUYER _--_--_--_--_--_--_--_--_--_--_
lajunta %>% 
  filter(is.na(Buyer))



# QUANTITY -__-__-__-__-__-__-__-__-__-__-
lajunta %>% 
  filter(is.na(Quantity))



# TYPE _--_--_--_--_--_--_--_--_--_--_
lajunta %>% 
  filter(is.na(Type))



# WEIGHT -__-__-__-__-__-__-__-__-__-__-
lajunta %>% 
  filter(is.na(Weight))



# PRICE _--_--_--_--_--_--_--_--_--_--_
lajunta %>% 
  filter(is.na(Price))



# REPROD -__-__-__-__-__-__-__-__-__-__-
# see any patterns?
lajunta %>% 
  dplyr::select(Type, Weight, Reprod, URL) %>% 
  filter(is.na(Reprod)) %>% 
  as.data.frame()

# You can see the distribution of missing Reprod values
lajunta %>% 
  filter(is.na(Reprod)) %>% 
  ggplot() +
  geom_col(mapping = aes(x = Weight, y = Price))

```
