---
title: "Collection (La Junta)"
author: "Connor Krenzer"
date: "3/11/2021"
output: html_document
---

# Introduction

I figure it's time to dig a bit deeper into Winter Livestock's market reports. I plan on checking all the market reports on the website to find historical data--but only for La Junta, CO. The old method works well enough, but limitations in available data using the old scripts make this approach far more lucrative.

This implementation assumes a few things about the structure of the text, so the scraper may need to be restricted to a certain date range.








# How It Works

To do this, I will first check whether the web scraper requires "\n" or "\r" as its delimiter. How do you do that? I'm glad you asked. I will do a strsplit() on the text data with "\n" as the delimiter. If the length of the resulting list is less than 30, I will try again using "\r" as the delimiter. If neither work, the algorithm moves onto the next market report. Most market reports are empty.


```{r setup-packages, include = FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      include = TRUE,
                      results = "hide",
                      error = F,
                      warning = F,
                      message = F)

# Packages
if(!require(pacman)) install.packages("pacman")
pacman::p_load(rvest, stringr, tidyr, readr, dplyr, lubridate)

```









# Remove File

If you would like to remove the csv file on your computer when updating data (meaning you would like to start from scratch), you can run the below chunk to remove the file.

```{r remove csv}

if(file.exists("La Junta Market Reports.csv")){
  file.remove("La Junta Market Reports.csv")
}

```










# Collection()

This function is a wrapper around the web scraper.


```{r collection()}

collection <- function(urls){
  
  
  for(URL in urls){
    
    # Reading in data from webpage ------------------------------
    # Saving the webpage into a variable
    webpage <- read_html(URL)
    
    # Saving data from the webpage written in html
    # The nodes are consistent across market reports
    livestock_data_html <- html_nodes(webpage, "div:nth-child(9) div.sml")
    
    # Converting to plain text
    livestock_data <- html_text(livestock_data_html)
    
    
    
    
    
    
    
    
    # Determining appropriate delimiter -------------------------
    # The first delimiter we will try
    delimiter <- "\n"
    
    # Asks whether the list has more than 30 entries after the strsplit(). If there is only one entry, then that means the format uses "\r" as its delimiter or does not contain information we care about.
    if(length(strsplit(livestock_data, delimiter)) < 30){
      
      # If the first delimiter does not get us what we are looking for, we will try the second delimiter instead.
      delimiter <- "\r" 
      
      # Now we test the second delimiter...
      if(length(strsplit(livestock_data, delimiter)) < 30){
        
        # If the second delimiter fails to deliver, skip to the next iteration of the loop (this webpage does not contain information we care about).
        next 
        
      }#end of nested if()
    }#end of if()
    
    
    
    
    
    
    
    
    
    # Determining location (La Junta) -------------------------
    # If we've made it this far, that means that there is information about a market report on the webpage.
    # We now need to determine whether this market report is for La Junta, CO.
    # This shouldn't be too hard. We can assume that "La Junta" will not appear in non-La Junta, CO market reports. Conversely, if we find "la junta" anywhere in the text, then we found a La Junta market report.
    
    
    # making all the text lowercase (to avoid the "la junta" != "La Junta" problem)
    livestock_data <- str_to_lower(livestock_data)
    
    # Now we can search for "la junta" in the text
    if(!str_detect(livestock_data, pattern = "la junta")){
      
      # If "la junta" is not found in the data, that means that this is not a La Junta market report--skip to the next iteration of the loop.
      next
    }
    
    
    
    
    
    
    
    
    
    
    # Finding the date of sale ----------------------------------------
    # Let's store the results from the strsplit() into a variable to more easily extract the date
    livestock_data <- strsplit(livestock_data, delimiter)
    
    
    # Saves the first entry that finds digits then the word "cattle"--this entry contains the sentence from which we can pull the date
    date_of_sale <- livestock_data[[1]][sapply(livestock_data, str_detect, "\\d+\\s*cattle") %>% 
                                          which() %>% 
                                          min()]
    # removing punctuation from the sentence
    date_of_sale <- str_remove_all(date_of_sale, "\\.|,|;")
    
    
    # extracting the month the sale took place from the sentence
    month_of_sale <- str_extract(date_of_sale, "january|february|march|april|may|june|july|august|september|october|november|december|jan|feb|mar|apr|jun|jul|aug|sept|oct|nov|dec")
    
    # the date of the sale in "month ##nd YYYY" format (Ex. "march 9th 2021")
    date_of_sale <- str_extract(date_of_sale, paste0(month_of_sale, "\\s+\\d{1,2}[a-z]{0,2}\\s+\\d{4}")) %>% 
      #replacing the month name with its corresponding number
      str_replace_all(month_of_sale,
                      case_when(month_of_sale == "january" ~ "1",
                                              month_of_sale == "february" ~ "2",
                                              month_of_sale == "march" ~ "3",
                                              month_of_sale == "april" ~ "4",
                                              month_of_sale == "may" ~ "5",
                                              month_of_sale == "june" ~ "6",
                                              month_of_sale == "july" ~ "7",
                                              month_of_sale == "august" ~ "8",
                                              month_of_sale == "september" ~ "9",
                                              month_of_sale == "october" ~ "10",
                                              month_of_sale == "november" ~ "11",
                                              month_of_sale == "december" ~ "12",
                                              month_of_sale == "jan" ~ "1",
                                              month_of_sale == "feb" ~ "2",
                                              month_of_sale == "mar" ~ "3",
                                              month_of_sale == "apr" ~ "4", #you don't need to repeat may, remember
                                              month_of_sale == "jun" ~ "6",
                                              month_of_sale == "jul" ~ "7",
                                              month_of_sale == "aug" ~ "8",
                                              month_of_sale == "sept" ~ "9",
                                              month_of_sale == "oct" ~ "10",
                                              month_of_sale == "nov" ~ "11",
                                              month_of_sale == "dec" ~ "12",
                      )) %>% 
      #removing letters and replacing space characters with hyphens
      str_replace_all("[a-z]", "") %>% 
      str_replace_all("\\s", "-")
    
    
    
    
    
    # Let's store everything back into a character vector
    livestock_data <- as.vector(livestock_data[[1]])
    
    
    
    
    
    
    
    
    
    
    
    
    # Header Removal with Keywords ------------------------------------------------------
    # A set of keywords designed to remove heading information
    # and other information we are not interested in observing
    keywords <- "\\s+sold|\\s+sale|\\s+monday|\\s+tuesday|\\s+wednesday|\\s+thursday|\\s+friday|\\s+saturday|\\s+sunday|\\s+receipts|\\s+through|\\s+mostly|\\s+winter|\\s+summer|\\s+spring|\\s+fall|\\s+autumn|\\s+is\\s+|\\s+next|\\s+quality|\\s+mostly|\\s+noon|\\s+early|\\s+stock|\\s+steady|\\s+test\\s+|\\s+offer|\\s+selection|\\s+week|\\s+annual|\\s+package|consigned|\\s*now\\s+|special\\s+|\\s+higher|calves\\s&\\syearlings\\s*$|\\s+am\\s+|\\s+pm\\s+|\\s+a.m.\\s+|\\s+p.m.\\s+|report[:]?\\s+|la\\s+junta,|\\s+co$|\\*$"
    
    
    # Removes headings and unrelated information from the data
    livestock_data <- livestock_data[!str_detect(livestock_data, keywords)]
    
    
    
    
    
    
    
    
    
    # Removing Unwanted Sections -------------------------------------------------------
    
    # We can pull out the sales information by removing lines we do not
    # care about. Since we know that the information we want is stored
    # in lines that are much shorter than the others, we can pull out
    # lines that have fewer characters than some optimal number.
    # I chose 60. In other words, I am keeping only those lines
    # (which are stored as elements in the vector) that contain
    # fewer than 60 characters.
    livestock_data <- livestock_data[-c(which((nchar(livestock_data) > 60)))]
    
    
    # The livestock data starts each day with a person's name and then has the quantity, type, weight, and price
    # if the person made more than one purchase, the line starts with "\n\t\t"--this is the reason
    # why we cannot make use of str_trim(). We need the "\n\t\t" to indicate whether
    # that element in the vector is really another purchase by the same buyer.
    
    # You can see what I am talking about by running
    # the code below. The "\n\t\t" is hidden, but
    # you can see the highlighted boxes at the
    # beginning of the lines containing the pattern:
    str_view(livestock_data, "\n\t\t")
    
    
    
    
    
    
    
    
    
    
    
    
    
    # Buyer names -----------------------------------------------------
    # Step 1. Names of Buyers:
    # Pulling out the indices that contain the buyer's name.
    buyers <- livestock_data[!str_detect(livestock_data, "\n\t\t")]
    
    # removing leading newline characters
    buyers <- str_trim(buyers)
    
    # Extracting the buyer's name--the name ends when
    # the first "\t" is encountered. You can see that
    # with this line:
    str_view(buyers, ".*?(?=\t)")
    
    # The names of the buyers:
    buyers <- str_extract(buyers, ".*?(?=\t)")
    
    # removing unneeded white space
    buyers <- str_trim(buyers)
    
    
    
    
    
    
    
    
    
    
    # Adding Buyer Names Back In ----------------------------------------------
    # This section uses ID numbers to identify the buyer for
    # each sale listed on the market report. It then adds
    # the buyer's name back onto the lines from which it
    # was omitted.
    
    # I will provide each buyer an ID number to determine
    # which purchases that buyer made. I am not crazy about
    # using loops, as I could probably write a function
    # to do this task for me, but it does everything we
    # need it to.
    current_ID <- 0
    for(i in 1:length(livestock_data)){
      if(str_detect(livestock_data[i], "\n\t\t")){
        # Only multiple purchases will have the "\n\t\t" so we can
        # use "\n\t\t" as a placeholder for the beginning of the string,
        # where the buyer's name goes, and make it look the same as the
        # lines where the name is present
        livestock_data[i] <- str_replace(livestock_data[i], "\n\t\t",  paste(buyers[current_ID], "\t", sep = ""))
        
      } else if(!str_detect(livestock_data[i], "\n\t\t")){
        # if the index lands on a new buyer, give them
        # the next ID number
        current_ID <- current_ID + 1 
      }
      
    }# end of for loop
    
    # removing white leading and trailing white space
    livestock_data <- str_trim(livestock_data)
    
    
    
    
    
    
    
    
    
    # Data frame ---------------------------------------------------
    # Making a data frame
    livestock_data <- tibble(livestock_data)
    
    
    # making new columns based off the sections
    # separated by "\t"
    livestock_data <- livestock_data %>% 
      separate(livestock_data, into = c("buyer", "quantity", "weight", "price"), sep = "\t")
    
    # Replaces the space separating the quantity and the type with a semicolon
    livestock_data$quantity <- str_replace(livestock_data$quantity, " ", ";")
    
    # Making quantity and type their own columns
    livestock_data <- livestock_data %>% 
      separate(quantity, into = c("quantity", "type"), sep = ";")
    
    
    # Making quantity, weight, and price numeric datatypes
    livestock_data[c(2, 4, 5)] <- sapply(livestock_data[c(2, 4, 5)], as.numeric)
    
    # Adding the date from the market report as a column
    
    
    
    # We are just about there! All that remains is removing the section
    # with NA's introduced:
    as.data.frame(livestock_data)
    
    # How do we fix that? Well, since we may want that information
    # for future use, we won't want to remove it entirely with a
    # keyword search, as was done in the beginning. What we can
    # remove observations with NA values to keep only the data
    # that matches our desired format:
    livestock_data <- na.omit(livestock_data)
    
    
    
    
    
    
    
    
    
    
    # Finishing Touches --------------------------------------------
    
    # We have all the data we need, though there are still a couple
    # problems with our output.
    #   1. We want a date column
    #   2. We want to remove the "s" at the end of the 
    #      type column to account for cases when there is only one
    #      cow bought by the buyer (Ex. "black cow" and "black cows"
    #      should both read "black cow")
    
    livestock_data <- mutate(livestock_data, "date" = date_of_sale, .before = 1)
    
    # Adding in the date
    #livestock_data <- mutate(livestock_data, "date" = Sys.Date(), .before = 1)
    
    
    # Removing the plural of the type--this is especially helpful for
    # use with group_by()
    livestock_data$type <- str_replace(livestock_data$type, "s$", "")
    
    
    
    # Writing to CSV ---------------------------------------
    # We are now ready to write the data to a file!
    # Just remember to add in the column names if
    # this file is not already on your computer
    write_csv(x = livestock_data,
              file = "La Junta Market Reports.csv",
              append = T,
              col_names = F)
    
  } # end of for loop


}#end of collection()

```











# Implementation




The first problem that must be solved is: how many market report IDs do you check? I'll try the first 15000. It doesn't look like they have gotten past 14000, so the extra thousand should provide ample room.

```{r urls}

# This just parses together 15,000 different possible market report IDs
urls <- paste0("http://www.winterlivestock.com/lajunta.php?reportID=", 1:15000, "#marketreport")

```



Now we run the function. This may take quite a bit of time, please remember...

```{r execute}

# No need to save the output--this function writes to a csv file...
collection(urls = urls)

```


Excellent. The next step is to read in the data we just collected. The reason I wrote it to a csv file first is because I did not know whether the result would be too large to save directly into memory. I doubt it, but you never know!

If you have trouble reading this csv file, click on Tools > Global Options > RMarkdown, then set 'Evaluate chunks in directory:' to Project.

```{r reading file}

lajunta <- read_csv("La Junta Market Reports.csv")

```




# Cleaning

The final step before we are ready for analysis is data cleaning. There are two primary goals in the cleaning script. First, we want to add column names to the csv and put entries in chronological order. The second goal is to organize cattle into a smaller number of distinct groups.

Goal 1 is a piece of cake, but goal 2 is easier said than done. You can analyze some of my judgement calls and decide for yourself what needs to be done to group the different types of livestock.

Do you think that prices should be adjusted for inflation at some point in the future?

```{r cleaning}

# Step 1: adding a header to the csv --------------------------
lajunta <- readr::read_csv("La Junta Market Reports.csv", 
                           col_names = F)

# Adding in the column names
lajunta <- lajunta %>% 
  rename("Date" = X1,
         "Buyer" = X2,
         "Quantity" = X3,
         "Type" = X4,
         "Weight" = X5,
         "Price" = X6)

# Putting the entries in chronological order,
# first changing the data type to a Date:
lajunta$Date <- lubridate::mdy(lajunta$Date)
lajunta <- arrange(lajunta, Date)


# Step 2: categorizing cattle ---------------------------------
# Making a column assigning cattle reproductive status
lajunta <- lajunta %>% 
  mutate(Reprod = str_extract(lajunta$Type, "hfr$|str$|bull$|cow$"))

# The bulk of the edits
lajunta$Type <- lajunta %>% 
  dplyr::select(Type) %>% 
  unlist() %>% 
  str_remove_all("\\sx.*$|\\s[^\\s]*$") %>%  # removing " x hfr", " hfr", " x str", etc.
  str_replace_all("angus", "ang") %>% 
  str_remove_all("sim-") %>%                      # removing the sim- from sim-angus
  str_replace_all("sim\\s+ang", "ang") %>%        # removing "sim " from "sim angus"
  str_replace_all("\\s+sim", " ang") %>%          # replacing " sim" with " ang"
  str_replace_all("limousin", "lim") %>% 
  str_replace_all("limo", "lim") %>%              # replacing "limo" with "lim"
  str_replace_all(".*\\s+lim|lim flex", "lim") %>%# combining all different types of limousins 
  str_replace_all("gelbvieh", "gel") %>% 
  str_replace_all("hereford", "here") %>% 
  str_replace_all("charolais", "char") %>% 
  str_replace_all("brahman", "brah") %>%
  str_remove_all("wf-|bwf-|rwf-") %>%             # removing face types from the data when there is other info
  str_remove_all("\\s*&") %>%                     # removing ampersands
  str_replace_all("\\s?blk\\s?", "black") %>%     # replacing abbreviation "blk" with "black"
  str_replace_all("red black", "black red") %>%   # reordering "red and black" to "black and red"
  str_replace_all(".*-.*", "mix") %>%             # replaces any named cross-breed with "mix"
  str_replace_all("balancer", "mix") %>%          # replaces "balancer" with "mix", since it's a hybrid
  str_remove_all("\\s+wf|\\s+bwf|\\s+rwf") %>%    # removing face types at after the name
  str_replace_all("^wf$", "bwf") %>%              # recoding "wf" as "bwf" because it performs similarly and has few observations
  str_remove_all("wf\\s+|bwf\\s+|rwf\\s+") %>%    # removing face types before the name
  str_replace_all("here|brah", "red") %>%         # replacing "here" and "brah" with "red" due to few observations and similar weight
  str_replace_all("black\\s+(ang)", "\\1") %>%    # removing color starts on this line
  str_replace_all("black\\s+(lim)", "\\1") %>% 
  str_replace_all("black\\s+(gel)", "\\1") %>% 
  str_replace_all("black\\s+(here)", "\\1") %>% 
  str_replace_all("black\\s+(char)", "\\1") %>% 
  str_replace_all("red\\s+(ang)", "\\1") %>% 
  str_replace_all("red\\s+(lim)", "\\1") %>% 
  str_replace_all("red\\s+(gel)", "\\1") %>% 
  str_replace_all("red\\s+(here)", "\\1") %>% 
  str_replace_all("red\\s+(char)", "\\1") %>%     # removing color ends on this line
  str_squish() %>% 
  str_trim()


#combining all colors into a new category--"clr" (optional)
lajunta$Type <- lajunta %>% 
  dplyr::select(Type) %>% 
  unlist() %>% 
  str_replace_all("black red|black|red|gray|grey", "clr") %>%   # "black red," "black," and "red" are all 
  str_replace_all("bwf|rwf|wf", "face") %>%
  str_replace_all("face", "clr")                  # remove this line if you want to stop pooling the face group with the color group

# We now have six categories in the "Type" variable




# Step 3: Writing to file -------------------------------------
write_csv(x = lajunta,
          file = "La Junta Market Reports.csv",
          append = F,
          col_names = T)


```








