---
title: "La Junta Predictions"
author: "Connor Krenzer"
date: "2/6/2021"
output: html_document
runtime: shiny
---

```{r setup-packages-data, include = FALSE}
# Don't worry about this, it's just to make the output
# look how I want it to...
knitr::opts_chunk$set(echo = FALSE,
                      results = "hide",
                      error = F,
                      warning = F,
                      message = F)


if(!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, lubridate, readr, shiny, patchwork)


# Reading in the file and removing the URL column
lajunta <- read_csv("https://raw.githubusercontent.com/Ckrenzer/Winter-Livestock-Data/main/La%20Junta%20Market%20Reports.csv",
                    col_types = cols(Date = "D",
                                     Buyer = col_factor(),
                                     Quantity = col_double(),
                                     Type = col_factor(),
                                     Weight = col_double(),
                                     Price = col_double(),
                                     URL = col_character(),
                                     Reprod = col_factor())) %>% 
  dplyr::select(-URL)

```

I want to bring your attention to the distinct types of cattle and how many of them appear in the data. See the data dictionary ("data KEY.txt") in the Collection folder for more info.

```{r plots, results = "show"}
# Shows the counts in each Reprod group
knitr::kable(lajunta %>% 
               count(Reprod) %>% 
               rename("Number of Appearances in Data" = n, "Reproductive Status" = Reprod) %>% 
               arrange(desc(`Number of Appearances in Data`)))

# Shows the counts of each Type
knitr::kable(lajunta %>% 
               count(Type) %>% 
               rename("Number of Appearances in Data" = n) %>% 
               arrange(desc(`Number of Appearances in Data`)))

# overall plot of weight vs price (after removing the outlier)
lajunta %>% 
  filter(Price < 475) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Weight, y = Price, color = Reprod)) +
  ggtitle("Weight vs. Price") +
  theme_dark()

# weight vs price by type (after removing the outlier)
lajunta %>%
  filter(Price < 475) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Weight, y = Price, color = Reprod)) +
  facet_wrap(~Type) +
  ggtitle("Weight vs. Price by Type") +
  theme_dark()

# weight vs price by reproductive status (after removing the outlier Reprod NA values)
lajunta %>%
  filter(!is.na(Reprod),
         Price < 475) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Weight, y = Price, color = Type)) +
  facet_wrap(~Reprod) +
  ggtitle("Weight vs. Price by Reproductive Status") +
  theme_dark()



# Price changes over time
q1 <- lajunta %>%  
  filter(Reprod %in% c("str", "hfr")) %>% 
  group_by(Date, Reprod) %>% 
  summarize(Price = median(Price)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Date, y = Price)) +
  geom_line(mapping = aes(x = Date, y = Price)) +
  facet_wrap(~Reprod) +
  ggtitle("Price Changes Over Time") +
  theme_dark()


q2 <-  lajunta %>%  
  filter(Reprod %in% c("bull", "cow")) %>% 
  group_by(Date, Reprod) %>% 
  summarize(Price = median(Price)) %>% 
  ggplot() +
  geom_point(mapping = aes(x = Date, y = Price)) +
  geom_line(mapping = aes(x = Date, y = Price)) +
  facet_wrap(~Reprod) +
  theme_dark()

# using patchwork's "/" operator
q1 / q2

```

Now, there's quite a bit that can be said for this information. On my end, it took about 8 hours to get down to six different types of cattle (and don't get me started on the hoops jumped through for the next section)--only to find out that there is basically no difference between the breeds. The reproductive status of the cattle, along with the weight, are what seem to drive the price. [Update--time spent fixing problems that pop up periodically accumulates and takes waaay longer than 8 hours!]

These prices are rather consistent, too. There are virtually no outliers. The bright side of this is that we can easily model the price. Since there are four distinct categories, four linear regression models seem appropriate.

```{r linear models}

steer_fit <- lm(data = lajunta, subset = Reprod == "str", formula = Price ~ Date + Quantity + poly(Weight, degree = 2))

heifer_fit <- lm(data = lajunta, subset = Reprod == "hfr", formula = Price ~ Date + Quantity + poly(Weight, degree = 2))

# a 12th degree polynomial gives us an R-squared of .54, but I think that is due to overfitting
cow_fit <- lm(data = lajunta, subset = Reprod == "cow", formula = Price ~ Date + poly(Weight, 3))

# a 6th degree polynomial gives an R-squared of .87, but I think that is due to overfitting
# plus, our sample is small so we want to keep the number of predictors small
bull_fit <- lm(data = lajunta, subset = Reprod == "bull", formula = Price ~ Date + poly(Weight, 2))

```


# Price Estimator
You can provide inputs below to estimate the price of the cattle you are selling. Prices for cows and bulls are very flat, so the model cannot identify price changes in these categories very well. Changes in the prices of cows and bulls are remarkably small, however, so this shouldn't really be a problem. You should be skeptical of the results for cows and bulls, however.

Another questionable decision I made was including the date as a predictor. OLS isn't well suited for variables pertaining to time, so (to reiterate) use this model with caution.

Compare the estimated price with the graphs above to evaluate the model.

### Describe Your Livestock
```{r str predictions, results = "show"}
ui <- fluidPage(
  
  # The four inputs we are interested in seeing
  selectInput("type", "What type of cattle are you interested in?", choices = c("steer", "heifer", "cow", "bull")),
  dateInput("date", "What is the date of the sale?"),
  numericInput("quantity", "How many are being sold?", 1, min = 1),
  numericInput("weight", "What is the livestock's weight in pounds?", 750, min = 25),
  
  # The output box, which uses data provided by the server
  verbatimTextOutput("price"),
  
)

server <- function(input, output, session) {
  
  type <- reactive({
    input$type
  })
  
  # the observe() function allows the input to get updated as the user changes the values around
  observe({
    #The if statements check the user's input for reproductive status,
    #then prints the model's prediction for the price.
    if(type() == "steer"){
      # the price, which gets referenced in the UI
      output$price <- renderText({
        paste0("ESTIMATED PRICE: ", "$", round(predict.lm(object = steer_fit, data.frame(Date = input$date, Quantity = input$quantity, Weight = input$weight)), 2))
        
      })
      
      
      
    } else if(type() == "heifer"){
      # the price, which is referenced in the UI
      output$price <- renderText({
        paste0("ESTIMATED PRICE: ", "$", round(predict.lm(object = heifer_fit, data.frame(Date = input$date, Quantity = input$quantity, Weight = input$weight)), 2))
        
      })
      
      
      
    } else if(type() == "cow"){
      # the price, which is referenced in the UI
      output$price <- renderText({
        paste0("ESTIMATED PRICE: ", "$", round(predict.lm(object = cow_fit, data.frame(Date = input$date, Quantity = input$quantity, Weight = input$weight)), 2))
        
      })
      
      
      
    } else {
      
      # the price, which is referenced in the UI
      output$price <- renderText({
        paste0("ESTIMATED PRICE: ", "$", round(predict.lm(object = bull_fit, data.frame(Date = input$date, Quantity = input$quantity, Weight = input$weight)), 2))
        
      })
    }#end of conditional
    
  })#end of observe()
  
}


shinyApp(ui, server)
```